---
title: "station_line_test_plot"
author: "Mallika Gupta"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Set up 

####Load Libraries
```{r}
#read in data 
library(readr)
#plotting 
library(ggplot2)
library(tidyverse)
#arranging plots in panel
library(ggpubr)
#creating interactive maps in R 
library(leaflet)
#splitting and averaging by unique values
library(data.table)
```

####Read in Files 
```{r}
#cast and bottle preprocessed data 
cast_bottle <- read_csv("data/bottle_and_cast.csv", show_col_types = FALSE)
cast_bottle

#descriptions of the variables in the cast and bottle preprocessed data 
descrip <- read_csv("data/Bottle Field Descriptions (1).csv", show_col_types = FALSE)
```


####Subset data set to look at particular station across quarters in 2000

```{r}
# cast_bottle$Sta_ID %>% substr(1, 5)=="093.3"  # determining how to select for line number within Station ID
# cast_bottle$Date == "01/07/2000"  # determining how to select for line number within Station ID
#  #Alternative subsetting method
# line_93.3 <- subset(cast_bottle, substr(cast_bottle$Sta_ID, 1, 5)=="093.3", select= c(Date, Depthm, O2ml_L))

# ## Quarter 1
# line_93.3_2000_1 <- cast_bottle[which(substr(cast_bottle$Sta_ID, 1, 5)=="093.3" & cast_bottle$Quarter == 1 & cast_bottle$Year == 2000),] %>%
#   mutate(Station = Sta_ID %>% substr(7, 11)) # Adding separate column with station number
# 
# # Data frame with all observations from line 93.3 in Quarter 1 of 2000 
# line_93.3_2000_1  
# 
# 
# ## Quarter 2
# line_93.3_2000_2 <- cast_bottle[which(substr(cast_bottle$Sta_ID, 1, 5)=="093.3" & cast_bottle$Quarter == 2 & cast_bottle$Year == 2000),] %>%
#   mutate(Station = Sta_ID %>% substr(7, 11)) # Adding separate column with station number
# 
# # Data frame with all observations from line 93.3 in Quarter 1 of 2000 
# line_93.3_2000_2  
# 
# 
# ## Quarter 3
# line_93.3_2000_3 <- cast_bottle[which(substr(cast_bottle$Sta_ID, 1, 5)=="093.3" & cast_bottle$Quarter == 3 & cast_bottle$Year == 2000),] %>%
#   mutate(Station = Sta_ID %>% substr(7, 11)) # Adding separate column with station number
# 
# # Data frame with all observations from line 93.3 in Quarter 1 of 2000 
# line_93.3_2000_3
# 
# 
# ## Quarter 4
# line_93.3_2000_4 <- cast_bottle[which(substr(cast_bottle$Sta_ID, 1, 5)=="093.3" & cast_bottle$Quarter == 4 & cast_bottle$Year == 2000),] %>%
#   mutate(Station = Sta_ID %>% substr(7, 11)) # Adding separate column with station number
# 
# # Data frame with all observations from line 93.3 in Quarter 1 of 2000 
# line_93.3_2000_4
```
###Creating Graphs 

```{r}
# Graphing oxygen concentration at different depths for each station line through different quarters of 2000
# station_93.3_2000_1_slice <- line_93.3_2000_1 %>%
#   ggplot(aes(x = Station, y = -Depthm)) +
#   geom_point(aes(color = O2ml_L), size = 3)
# station_93.3_2000_1_slice
# 
# station_93.3_2000_2_slice <- line_93.3_2000_2 %>%
#   ggplot(aes(x = Station, y = -Depthm)) +
#   geom_point(aes(color = O2ml_L), size = 3)
# station_93.3_2000_2_slice
# 
# station_93.3_2000_3_slice <- line_93.3_2000_3 %>%
#   ggplot(aes(x = Station, y = -Depthm)) +
#   geom_point(aes(color = O2ml_L), size = 3)
# station_93.3_2000_3_slice
# 
# station_93.3_2000_4_slice <- line_93.3_2000_4 %>%
#   ggplot(aes(x = Station, y = -Depthm)) +
#   geom_point(aes(color = O2ml_L), size = 3)
# station_93.3_2000_4_slice
```


```{r, fig.width = 8, fig.height = 10}
# Cleaning up code by using facet_wrap
line_93.3_2000_slices <- cast_bottle[which(substr(cast_bottle$Sta_ID, 1, 5)=="093.3" & cast_bottle$Year == 2000),] %>%
  mutate(Station = Sta_ID %>% substr(7, 11)) %>% # Adding separate column with station number 
  pivot_longer(cols = Quarter) %>%
  ggplot(aes(x = Station, y = -Depthm)) +
  geom_point(aes(color = O2ml_L), size = 3) +
  facet_wrap(~ value, nrow = 4, scales = "free_x") +
  labs(x = "Station", y = "Depth (m)") 

line_93.3_2000_slices
```


# Questions
Above graphs bring up question about our data: why are there fewer stations sampled alone line 93.3 in the 3rd quarter than the others? 
  - Also seems like the granularity of depth was much higher in quarter 2 than the others (can tell from table dimensions). 
Maybe we consider using "Distance - Nautical miles from coast intercept, calculated from estimated latitude and longitude" instead of station line here so we can see a scaled representation of how this points are distanced from the coastline
  - Could bring up problems in representation since latitude and longitude aren't exactly on top of the station line (but could also make the visualization more accurate + sacirficing the straight lines along station numbers isn't that bad)
  
  
Add some transformation of the y-axis?
Color scale can be also be subjected to transformation to bring out detail in the different ranges of oxygen - could also be adjusted by the thresholds
Flip the color scale to have light colors for low oxygen concentration
There are a lot of default color scales: sequential, diverging, 3 color sequential ones etc.
Scales package could allow custon transformations to be used in ggplot
Check: ggplot2 - scale_fill_gradient or scale_color_gradient?

Diverging gradient around the threshold - red values for under the threshold and diverges again for higher values of oxygen


```{r,  fig.width = 8, fig.height = 10}
#line: 	076.7
# Cleaning up code by using facet_wrap
line_076.7_2000_slices <- cast_bottle[which(substr(cast_bottle$Sta_ID, 1, 5)=="076.7" & cast_bottle$Year == 2000),] %>%
  mutate(Station = Sta_ID %>% substr(7, 11)) %>% # Adding separate column with station number 
  pivot_longer(cols = Quarter) %>%
  ggplot(aes(x = Station, y = -Depthm)) +
  geom_point(aes(color = O2ml_L), size = 3) +
  facet_wrap(~ value, nrow = 4, scales = "free_x") +
  labs(x = "Station", y = "Depth (m)")

line_076.7_2000_slices
```










